# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Load Data located in forked repository

```{r Reading Activity Data}
activity <- read.csv("activity.csv")
# Check data and variable types
summary(activity)
str(activity)
```

Aggregate data for analysis as Steps taken per day and check output

```{r Aggregate steps per day}
steps_day <- aggregate(steps~date, activity, sum)
head(steps_day)
summary(steps_day)
```

Aggregate data for analysis of steps taken per daily time interval and check output

```{r Aggregate steps per interval}
steps_interval <- aggregate(steps~interval, activity, mean)
summary(steps_interval)
```


## What is mean total number of steps taken per day?
plot histogram of steps per day
```{r}
library(ggplot2)
ggplot(steps_day, aes(x=steps)) + geom_histogram()
```

Find the mean of the total daily steps.

```{r}
mean(steps_day$steps, na.rm = TRUE)
```

Find the median of the total daily steps

```{r}
median(steps_day$steps, na.rm = TRUE)
```
These values agree with the Summary statistics of the steps_day aggregate from above.

## What is the average daily activity pattern?

plot time series of steps taken per interval over total time frame
```{r}
plot(steps_interval$steps~steps_interval$interval, type = "l")
```

Find the time interval which contains the maximum number of steps, on average

```{r}
maxInterval <- steps_interval[steps_interval$steps == max(steps_interval$steps),]
maxInterval
```
The 5 minute interval with the most steps over the time frame studied 
appears to be at 8:35 in the morning

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
